---
title: "Integration and Differential Expression of mice TA muscle data at 5m/o and 24 m/o"
output: html_document
date: "2022-08-06"
---

## Integration and Differential Expression of mice TA muscle data at 5m/o and 24 m/o

In this document, we have included the script used to integrate the 5m/o TA muscle data and 24m/o muscle datasets. We have also identified differentially expressed gene signatures in different cell_types.

### Loading the necessary packages

```{r, warning=FALSE, message=FALSE}
library(Seurat)
library(dplyr)
library(ggplot2)
library(cowplot)
library(xlsx)
```

### Loading the muscle Seurat object from pre-processing
```{r, warning=FALSE, message=FALSE}
fivemonth_muscle <- readRDS("fivemonth_muscle.RDS")
twentyfourmonth_muscle <- readRDS("twentyfourmonth_muscle.RDS")
```

### integrating the muscle cells pre-processed data from 5m/o and 24 m/o mice
```{r,warning=FALSE, message=FALSE }
fivemonth_muscle@meta.data[["orig.ident"]] <- "A"
twentyfourmonth_muscle@meta.data[["orig.ident"]] <- "B"

integration.list.muscle <- list(fivemonth_muscle, twentyfourmonth_muscle)
for (i in 1:length(integration.list.muscle)) {
  integration.list.muscle[[i]] <- NormalizeData(integration.list.muscle[[i]], verbose = FALSE)
  integration.list.muscle[[i]] <- FindVariableFeatures(integration.list.muscle[[i]], selection.method = "vst", 
                                                       nfeatures = 2000, verbose = FALSE)
}

integration.anchors.muscle <- FindIntegrationAnchors(object.list = integration.list.muscle, dims = 1:30, verbose = FALSE)
integration.integrated.muscle <- IntegrateData(anchorset = integration.anchors.muscle, dims = 1:30)
integration.integrated.muscle <- ScaleData(integration.integrated.muscle, verbose = FALSE)
integration.integrated.muscle <- RunPCA(integration.integrated.muscle, npcs = 30, verbose = FALSE)
integration.integrated.muscle <- RunUMAP(integration.integrated.muscle, reduction = "pca", dims = 1:30)
DimPlot(integration.integrated.muscle, reduction = "umap", label = TRUE)
DimPlot(integration.integrated.muscle, reduction = "umap", group.by = "orig.ident", label = TRUE, repel = TRUE) + NoLegend()

integration.integrated.muscle <- ScaleData(integration.integrated.muscle, verbose = FALSE)
muscle_markers <- FindAllMarkers(integration.integrated.muscle)
A <- subset(integration.integrated.muscle, orig.ident == "A")
B <- subset(integration.integrated.muscle, orig.ident == "B")

A_cells <- rownames(A)
B_cells <- rownames(B)

A_markers <- FindMarkers(integration.integrated.muscle, ident.1 = "A", ident.2 = NULL, group.by = "orig.ident", only.pos = TRUE)
B_markers <- FindMarkers(integration.integrated.muscle, ident.1 = "B", ident.2 = "A", group.by = "orig.ident", only.pos = TRUE)

top10 <- muscle_markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)

write.xlsx(A_markers, "5month positive markers.xlsx")
write.xlsx(B_markers, "24month positive markers.xlsx")
```

## Differential Expression of genes in different cell-type of mice at different age (5m/o and 24m/o mice)
```{r,warning=FALSE, message=FALSE }
Idents(integration.integrated.muscle)
diffexp5month <- subset(integration.integrated.muscle, orig.ident == "A")
diffexp24month <- subset(integration.integrated.muscle, orig.ident == "B")

diffexp_5m_type2 <- FindMarkers(diffexp5month, ident.1 = "Type IIb Myonuclei", ident.2 = "Type IIx Myonuclei", verbose = FALSE)
diffexp_24m_type2 <- FindMarkers(diffexp24month, ident.1 = "Type IIb Myonuclei", ident.2 = "Type IIx Myonuclei", verbose = FALSE)
diffexp_24m_type2.2 <- FindMarkers(diffexp24month, ident.1 = "Type IIb Myonuclei #2", ident.2 = "Type IIx Myonuclei #2", verbose = FALSE)
diffexp_24m_type2b <- FindMarkers(diffexp24month, ident.1 = "Type IIb Myonuclei", ident.2 = "Type IIb Myonuclei #2", verbose = FALSE)
diffexp_24m_type2x <- FindMarkers(diffexp24month, ident.1 = "Type IIx Myonuclei", ident.2 = "Type IIx Myonuclei #2", verbose = FALSE)


# differential expression between Satellite vs myocyte
diffexp_5m_sattype2x <- FindMarkers(diffexp5month, ident.1 = "Satellite Cells", ident.2 = "Type IIx Myonuclei", verbose = FALSE)
diffexp_5m_sattype2b <- FindMarkers(diffexp5month, ident.1 = "Satellite Cells", ident.2 = "Type IIb Myonuclei", verbose = FALSE)
diffexp_24m_sattype2b <- FindMarkers(diffexp24month, ident.1 = "Satellite Cells", ident.2 = "Type IIb Myonuclei", verbose = FALSE)
diffexp_24m_sattype2x <- FindMarkers(diffexp24month, ident.1 = "Satellite Cells", ident.2 = "Type IIx Myonuclei", verbose = FALSE)
diffexp_24m_sattype2b.2 <- FindMarkers(diffexp24month, ident.1 = "Satellite Cells", ident.2 = "Type IIb Myonuclei #2", verbose = FALSE)
diffexp_24m_sattype2x.2 <- FindMarkers(diffexp24month, ident.1 = "Satellite Cells", ident.2 = "Type IIx Myonuclei #2", verbose = FALSE)

gene.names <- c()
gene.names <- c(gene.names, rownames(diffexp_5m_type2), rownames(diffexp_24m_type2), rownames(diffexp_24m_type2.2))
gene.names <- c(gene.names, rownames(diffexp_24m_type2b), rownames(diffexp_24m_type2x), rownames(diffexp_5m_sattype2x))
gene.names <- c(gene.names, rownames(diffexp_5m_sattype2b), rownames(diffexp_24m_sattype2b), rownames(diffexp_24m_sattype2x))

gene.names <- c(gene.names, rownames(diffexp_24m_sattype2b.2), rownames(diffexp_24m_sattype2x.2))
unique.gene.names <- unique(gene.names)
diffexp.circgene.celltype <- intersect(unique.gene.names, unique(df_clean2$Gene)) 
```

## differential expression by timepoints
```{r, warning=FALSE, message=FALSE}
# differential expression by timepoints
integration.integrated.muscle$celltype.age <- paste(Idents(integration.integrated.muscle), integration.integrated.muscle$orig.ident, sep = "_")
integration.integrated.muscle$celltype <- Idents(integration.integrated.muscle)
Idents(integration.integrated.muscle) <- "celltype.age"
typeIIb.age <- FindMarkers(integration.integrated.muscle, ident.1 = "Type IIb Myonuclei_A", ident.2 = "Type IIb Myonuclei_B", verbose = FALSE)
typeIIx.age <- FindMarkers(integration.integrated.muscle, ident.1 = "Type IIx Myonuclei_A", ident.2 = "Type IIx Myonuclei_B", verbose = FALSE)
satellitte.age <- FindMarkers(integration.integrated.muscle, ident.1 = "Satellite Cells_A", ident.2 = "Satellite Cells_B", verbose = FALSE)
myotendinous.age <- FindMarkers(integration.integrated.muscle, ident.1 = "Myotendinous Junction_A", ident.2 = "Myotendinous Junction_B", verbose = FALSE)
neuromuscular.age <- FindMarkers(integration.integrated.muscle, ident.1 = "Neuromuscular Junction_A", ident.2 = "Neuromuscular Junction_B", verbose = FALSE)


gene.names.age <- c(rownames(typeIIb.age), rownames(typeIIx.age), rownames(myotendinous.age), rownames(neuromuscular.age))
diffexp.circgene.age <- intersect(unique(gene.names.age), unique(df_clean2$Gene)) 
diffexp.circgene.celltype.age <- intersect(diffexp.circgene.age, diffexp.circgene.celltype)
```